Optimal strategies of influence in structured populations

### 2015-07-15 report/report.tex
updated reports and paper

Each individual, $i$, holds an opinion $S_i \in \{ -1,1\}$ that is observable by others in its local neighborhood and is correct when $S_i = \mbox{sign}(G)$. Individuals update their states based on social and personal information and a Bayesian decision rule \cite{nitzan1982optimal,perez2011collective, Perreault, torney2015} is employed. This means that an individual will switch to the state that is the most likely to be correct, given all available information. Formally, individual $i$ will adopt the state $S_i=1$ if
\begin{equation}
 P\left(G \ge 0\big| g_i, \mathbf{S}_{j\in \mathcal{N}_i}\right)  > P\left(G < 0\big| g_i, \mathbf{S}_{j\in \mathcal{N}_i}\right)
\label{opt_sol1}
\end{equation}
where $\mathcal{N}_i$ defines the social interaction neighborhood. To impose a decision rule as a function of $g_i$ we rearrange Eqn.~\ref{opt_sol1} using Bayes' rule, assume individuals place equal weighting on the opinions of neighbors, that they know the level of noise in the environmental cue, and that it is symmetric around zero. Eqn.~\ref{opt_sol1} may then be written as,
\begin{equation}
\frac{\exp{ \left( - \frac{\vert g_i -  \infty \vert}{\sigma} \right)}}{\exp{ \left( - \frac{\vert g_i + \infty \vert}{\sigma} \right)}} > \chi^{1-2x_L}.
\label{rule}
\end{equation}
Here we have denoted the fraction of neighbors an individual observes in the state $S=1$ as $x_L$, used $\pm \infty$ as estimates of the actual value of $G$ (these will cancel), and introduced a social weighting variable $\chi$ which may be interpreted as the belief that a single neighbor is correct rescaled by the neighborhood size. This rescaling means decisions are based on the fraction of neighbors that are correct in the local neighborhood and not the absolute number, ensuring a comparison can be made between social networks with varying average degree.

Rearranging Eqn.~\ref{rule} leads to a decision rule such that individual $i$ will adopt the opinion $S_i = 1$ iff
\begin{equation}
g_i >  \sigma\ln\chi  \left(\frac{1}{2}-x_L\right) 
\label{opt_rule}
\end{equation}

%, i.e. $\chi = P(S_j=\mbox{sign}(G))^{\vert \mathcal{N}_i \vert }$  the and denote the fraction of neighbors an individual observes in the state $S=1$ as $x_L$. Then the decision rule for the focal individual is to switch to the state $1$ if

@@

We aim to assess optimal strategies assuming some cost associated with reducing the noise. Cost is per unit time and assumed to scale exponentially so that forcing $\sigma \rightarrow 0$ incurs exponentially increasing costs,
\begin{equation}
\mbox{Cost} = \tau (\sigma) e^{\beta N/\sigma}
\label{cost}
\end{equation}
Minimizing Eqn.~\ref{cost} means the optimal level of $\sigma$ is one for which
\begin{equation}
N^{-1} \frac{d \ln \tau}{d \sigma} = \beta \sigma^{-2}
\end{equation}
\begin{figure*}
\centering
\includegraphics[scale=0.4]{figures/fig-chi.eps} 
\caption{Probability to transition as a function of local neighbourhood for different values of the sociality weighting $\chi$}
\label{fig1}
\end{figure*}
We need to find $\tau$.
\section{Mean first passage time away from incorrect state}
Consider the process as a one-step Markov chain on the integer line from $0$ to $N$. A transition from state $j$ to $j+1$ is written $t^+(j)$ and an opposite transition is denoted $t^-(j)$. An exact expression for the mean first passage time from zero individuals correct to $N_{ESC}$ is \cite{gardiner}

@@

\section{Transition rates}
The dynamics of an update step in stochastic decision process are as follows; first an individual is selected at random from the population, this individual has a personal estimate of the environmental cue $g_i$ and observes the opinions of all its neighbors. The decision rule of Eqn.~\ref{opt_rule} is then employed and the individual will adopt, or remain in, the state most likely to be correct.
For any value of social information $x_L$ there is a value of personal information above which the individual will switch to $S=1$. We denote this value as $g_c$ and from Eqn.~\ref{opt_rule} 
\begin{equation}
g_c(x_L) = \sigma\ln\chi  \left(\frac{1}{2}-x_L\right) 
\end{equation}
If $g_i>g_c$ then the individual will switch to $S=1$ otherwise $S=-1$. 
Dealing first with the transition rate toward the correct state, we may write this as
\begin{equation}
%t^+(x) = P(S_i=-1)\sum_{j=0}^{N_s}P\left(g_i>g_c(j/N_s)\right)P\left(x_L=j/N_s|x\right)
t^+(x) = P(S_i=-1)\sum_{x_{Li}}P\left(g_i>g_c(x_{Li})\right)P\left(x_{Li}|x\right)
\end{equation}
where $i$ denotes the selected individual and the summation is taken over all possible values of $x_{Li}$.
he product of the probability that the focal individual is in the state $S=-1$ by the probability of a given value of $x_L$ by the

>>>>

Each individual, $i$, holds an opinion $S_i \in \{ -1,1\}$ that is observable by others in its local neighborhood and is correct when $S_i = \mbox{sign}(G)$. Individuals update their states based on social and personal information and a Bayesian decision rule \cite{nitzan1982optimal,perez2011collective, Perreault, torney2015} is employed. This means that an individual will switch to the state that is the most likely to be correct, given all available information. Formally, individual $i$ will adopt the state $S_i=1$ if
\begin{equation}
 P\left(G \ge 0\big| g_i, \mathbf{S}_{j\in \mathcal{N}_i}\right)  > P\left(G < 0\big| g_i, \mathbf{S}_{j\in \mathcal{N}_i}\right)
\label{opt_sol1}
\end{equation}
where $\mathcal{N}_i$ defines the social interaction neighborhood. To impose a decision rule as a function of $g_i$ we rearrange Eqn.~\ref{opt_sol1} using Bayes' rule, assume individuals place equal weighting on the opinions of neighbors, that they know the level of noise in the environmental cue, and that it is symmetric around zero. Eqn.~\ref{opt_sol1} may then be written as,
\begin{equation}
\frac{\exp{ \left( - \frac{\vert g_i -  \infty \vert}{\sigma} \right)}}{\exp{ \left( - \frac{\vert g_i + \infty \vert}{\sigma} \right)}} > \omega^{1-2x_L}.
\label{rule}
\end{equation}
Here we have denoted the fraction of neighbors an individual observes in the state $S=1$ as $x_L$, used $\pm \infty$ as estimates of the actual value of $G$ (these will cancel), and introduced a social weighting variable $\omega$ which may be interpreted as the belief that a single neighbor is correct rescaled by the neighborhood size. This rescaling means decisions are based on the fraction of neighbors that are correct in the local neighborhood and not the absolute number, ensuring a comparison can be made between social networks with varying average degree.

Rearranging Eqn.~\ref{rule} leads to a decision rule such that individual $i$ will adopt the opinion $S_i = 1$ iff
\begin{equation}
g_i >  \sigma\ln\omega  \left(\frac{1}{2}-x_L\right) 
\label{opt_rule}
\end{equation}

%, i.e. $\omega = P(S_j=\mbox{sign}(G))^{\vert \mathcal{N}_i \vert }$  the and denote the fraction of neighbors an individual observes in the state $S=1$ as $x_L$. Then the decision rule for the focal individual is to switch to the state $1$ if

@@

We aim to assess optimal strategies assuming some cost associated with reducing the noise. Cost is per unit time and assumed to scale exponentially so that forcing $\sigma \rightarrow 0$ incurs exponentially increasing costs,
\begin{equation}
\mbox{Cost} = \tau (\sigma) e^{\beta N/\sigma}
\label{cost}
\end{equation}
Minimizing Eqn.~\ref{cost} means the optimal level of $\sigma$ is one for which
\begin{equation}
N^{-1} \frac{d \ln \tau}{d \sigma} = \beta \sigma^{-2}
\end{equation}
\begin{figure*}
\centering
\includegraphics[scale=0.4]{figures/fig-chi.eps} 
\caption{Probability to transition as a function of local neighbourhood for different values of the sociality weighting $\omega$}
\label{fig1}
\end{figure*}
We need to find $\tau$.
\section{Mean first passage time away from incorrect state}
Consider the process as a one-step Markov chain on the integer line from $0$ to $N$. A transition from state $j$ to $j+1$ is written $t^+(j)$ and an opposite transition is denoted $t^-(j)$. An exact expression for the mean first passage time from zero individuals correct to $N_{ESC}$ is \cite{gardiner}

@@

\section{Transition rates}
The dynamics of an update step in stochastic decision process are as follows; first an individual is selected at random from the population, this individual has a personal estimate of the environmental cue $g_i$ and observes the opinions of all its neighbors. The decision rule of Eqn.~\ref{opt_rule} is then employed and the individual will adopt, or remain in, the state most likely to be correct.
For any value of social information $x_L$ there is a value of personal information above which the individual will switch to $S=1$. We denote this value as $g_c$ and from Eqn.~\ref{opt_rule} 
\begin{equation}
g_c(x_L) = \sigma\ln\omega  \left(\frac{1}{2}-x_L\right) 
\end{equation}
If $g_i>g_c$ then the individual will switch to $S=1$ otherwise $S=-1$. 
Dealing first with the transition rate toward the correct state, we may write this as
\begin{equation}
%t^+(x) = P(S_i=-1)\sum_{j=0}^{N_s}P\left(g_i>g_c(j/N_s)\right)P\left(x_L=j/N_s|x\right)
t^+(x) = P(S_i=-1)\sum_{x_{Li}}P\left(g_i>g_c(x_{Li})\right)P\left(x_{Li}|x\right)
\end{equation}
where $i$ denotes the selected individual and the summation is taken over all possible values of $x_{Li}$.
he product of the probability that the focal individual is in the state $S=-1$ by the probability of a given value of $x_L$ by the
